### 2021-07-03 rad-icfp-2021/correctness.tex
minor

\begin{tabular}{l c r}
   $\seml$!var!$_{T;i} \semr$ &=& \\
   $\seml$!op!$_{T;n} \semr$ &=& \\
   $\seml$!pair!$_{T;A\times B} \semr$ &=& \\
   $\seml$!proj!$_{T1;T2;T3} \semr$ &=& \\
   $\seml$!e!$\comp$!e!$\semr$  &=& \\
   $\seml$!e!$\pcomp$!e!$\semr$ &=& \\
   $\seml$!map2!$_{T;x,y.e}\semr$  &=& \\
   $\seml$!reduce!$_{T;x,y.e;e}\semr$  &=& \\
\end{tabular}

@@

\begin{tabular}{l c r}
    $\seml$!var!$_{T;i} \semr$ &=& \\
    $\seml$!op!$_{T;n} \semr$ &=& \\
    $\seml$!pair!$_{T;A\times B} \semr$ &=& \\
    $\seml$!proj!$_{T1;T2;T3} \semr$ &=& \\
    $\seml$!e!$\comp$!e!$\semr$  &=& \\
    $\seml$!e!$\pcomp$!e!$\semr$ &=& \\
    $\seml$!map2!$_{T;x,y.e}\semr$  &=& \\
    $\seml$!reduce!$_{T;x,y.e;e}\semr$  &=& \\
 \end{tabular}
\MH{Do I need any kind of proposition saying that this makes sense?
Might need to define on types as well to be clearer: It's an identity on type functor...}

>>>>

\begin{tabular}{l c l}
   $\seml$!var!$_{T;i} \semr$ &=& TODO \\
   $\seml$!op!$_{T;n} \semr$ &=& \\
   $\seml$!pair!$_{T;A\times B} \semr$ &=& \\
   $\seml$!proj!$_{T1;T2;T3} \semr$ &=& \\
   $\seml$!e!$\comp$!e!$\semr$  &=& \\
   $\seml$!e!$\pcomp$!e!$\semr$ &=& \\
   $\seml$!map2!$_{T;x,y.e}\semr$  &=& \\
   $\seml$!reduce!$_{T;x,y.e;e}\semr$  &=& \\
\end{tabular}

@@

\begin{tabular}{l c l}
    $\seml$!var!$_{T;i} \semr$ &=& TODO \\
    $\seml$!op!$_{T;n} \semr$ &=& \\
    $\seml$!pair!$_{T;A\times B} \semr$ &=& \\
    $\seml$!proj!$_{T1;T2;T3} \semr$ &=& \\
    $\seml$!e!$\comp$!e!$\semr$  &=& \\
    $\seml$!e!$\pcomp$!e!$\semr$ &=& \\
    $\seml$!map2!$_{T;x,y.e}\semr$  &=& \\
    $\seml$!reduce!$_{T;x,y.e;e}\semr$  &=& \\
 \end{tabular}
\MH{Do I need any kind of proposition saying that this makes sense?
Might need to define on types as well to be clearer: It's an identity on type functor...}

### 2021-07-03 rad-icfp-2021/relwork.tex
minor

\noindent \textbf{Cost function and memory accesses}
Our cost model \ref{sub:costModel} is simple but has limits. 
In particular, it does not capture well cost in term of
parallel computing, and memory management (cache optimisation is one of best way to optimise such linear algebra intensive computations).
For instance, in the earlier days of automatic differentiation, 
advanced checkpointing techniques were used to avoid storing all the intermediate values just to save memory 
and try avoiding saturation of RAM. 
These problems are harder to analyse and eventually it seems to come down to a lot of experimenting and benchmarking.

\noindent \textbf{Parallelism and GPU}
We tried to locally preserve parallelism as much as possible, 
and having a pure transformation should help with found parallelism. 
It would be interesting to implement on transformations and see how they perform on GPU.
It is well-kwown that the addition of references provoked by the standard reverse-mode transformation
restricts a lot of optimisations and it is not hard to find examples in C where the compiler does not detect simple optimisations 
because of the references introduced by reverse-mode.

@@

\begin{itemize}
    \item fancier types and datatypes (eg dependent, adt) 
    \item use in Probabilistic programming
    \item other optimisations like implicit function theorem
\end{itemize}

% \MH{useful links (Swift is close to our project, and the doc is very readable):}
% \begin{verbatim}https://docs.google.com/document/d/1_BirmTqdotglwNTOcYAW-ib6mx_jl-gH9Dbg4WmHZh0/edit# \end{verbatim}
% \begin{verbatim}
% 	https://github.com/apple/swift/blob/master/docs/DifferentiableProgramming.md#approaches-to-automatic-differentiation
% \end{verbatim}

@@

\MH{should talk a bit about very recent reverse AD in Dex \cite{paszke2021getting}}
\clearpage

>>>>

\noindent \textbf{Cost function and memory accesses}
Our cost model \ref{sub:costModel} is simple but has limits. 
In particular, it does not capture well cost in term of
parallel computing, and memory management (cache optimization is one of best way to optimize such linear algebra intensive computations).
For instance, in the earlier days of automatic differentiation, 
advanced checkpointing techniques were used to avoid storing all the intermediate values just to save memory 
and try avoiding saturation of RAM. 
These problems are harder to analyze and, eventually, it seems to come down to a lot of experimenting and benchmarking.

\noindent \textbf{Parallelism and GPU}
We tried to locally preserve parallelism. Having a pure transformation should help with found parallelism. 
It would be interesting to implement on transformations and see how they perform on GPU.
It is well-known that the addition of references provoked by the standard reverse-mode transformation
restricts a lot of optimizations and it is not hard to find examples in C where the compiler does not detect simple optimisations 
because of the references introduced by reverse-mode.

@@

\noindent \textbf{Further applications}
Reverse-mode differentiation is notoriously difficult to grasp, implement, and optimize.
We hope this work could help develop a new perspective on pure reverse-mode automatic differentiation.
Other applications include the use of fancier datatypes e.g. dependent types, 
in probabilistic programming, e.g. for HMC or Variational Inference. 
It is also of interest to see how to extend this work with other techniques which 
have had renewed interest such as implicit differentiation \cite{blondel2021efficient,lorraine2020optimizing}.

% 	https://github.com/apple/swift/blob/master/docs/DifferentiableProgramming.md#approaches-to-automatic-differentiation

@@

\MH{should talk a bit about the recent paper: reverse AD in Dex \cite{paszke2021getting}}
\clearpage
